---
title: "Lab 5"
author: "Ahmed Hassan"
date: "11/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initialization, echo=FALSE, include=FALSE}
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(readxl))
suppressPackageStartupMessages(require(tidyverse))
suppressPackageStartupMessages(require(lubridate))
suppressPackageStartupMessages(require(ggrepel))
library(ggplot2)
library(readxl)
library(dplyr)
library(tidyverse)
library(lubridate)
library(ggrepel)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

pers  <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

  
```

1)
```{r}
acc %>% group_by(DAY_WEEK) %>% 
  summarise(total = length(DAY_WEEK))
```
```
We can see based on the Summary that the 1st, 6th and 7th day of the week has the most accidents. With the 7t day of the week having the most. 
```

2)
```{r}
hurt <- pers %>% 
  filter(INJ_SEV == 4)

```

3)Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)

```{r}
col <- pers[, c("INJ_SEV", "MAKE", "STATE", "COUNTY")]
danger <- col %>%
  filter(INJ_SEV == 4) %>%
  group_by(STATE, MAKE) %>%
  summarize(n = n()) %>%
  group_by(STATE) %>%
  filter(n == max(n)) %>%
  filter(is.na(MAKE) == FALSE) %>%
  arrange(desc(n))
danger

```

4)Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)
```{r}

codes<- readxl::read_xlsx("codes.xlsx")
codes <- codes[, c( "State Code", "County Code", "State Name","County Name")] %>%
  rename("STATE" = "State Code") %>% 
  rename("COUNTY" = "County Code")

##codes
codes$STATE <- as.numeric(codes$STATE)
codes$COUNTY <- as.numeric(codes$COUNTY)

#Main DataFrames for Map
spatialDF <- acc[,c("STATE","COUNTY","LATITUDE", "LONGITUD")]
masterDF<-full_join(codes, spatialDF, by = "COUNTY","STATE")
masterDF<-masterDF %>% rename("STATE" = "STATE.x")

spatialDF$STATE <- as.numeric(spatialDF$STATE)
spatialDF$COUNTY <- as.numeric(spatialDF$COUNTY)


tempDF<-col %>% 
  group_by(STATE,COUNTY, MAKE) %>% 
  summarize(n = n()) %>% 
   group_by(STATE) %>%
  filter(is.na(MAKE) == FALSE) %>%
  filter(n == max(n)) %>% 
  arrange(desc(n))

masterDF<-full_join(masterDF, tempDF,by = "COUNTY","STATE")
masterDF %>% drop_na()
#Choropleth Map

ggplot(masterDF, aes(x=masterDF$LONGITUD, y=masterDF$LATITUDE))+geom_polygon() #Check on how to get just Outlines for states, or need point for every county/city --> inner states.
```



