---
title: "Lab 5"
author: "Ahmed Hassan"
date: "11/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r initialization, echo=FALSE, include=FALSE}
suppressPackageStartupMessages(require(ggplot2))
suppressPackageStartupMessages(require(dplyr))
suppressPackageStartupMessages(require(readxl))
suppressPackageStartupMessages(require(tidyverse))
suppressPackageStartupMessages(require(lubridate))
suppressPackageStartupMessages(require(ggrepel))
library(ggplot2)
library(readxl)
library(dplyr)
library(tidyverse)
library(lubridate)
library(ggrepel)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)

pers  <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)

  
```

1)
```{r}
acc %>% group_by(DAY_WEEK) %>% 
  summarise(total = length(DAY_WEEK))
```
```
We can see based on the Summary that the 1st, 6th and 7th day of the week has the most accidents. With the 7t day of the week having the most. 
```

2)
```{r}
hurt <- pers %>% 
  filter(INJ_SEV == 4)

```

3)Create a data frame containing the most dangerous vehicle make in each state. The number of persons fatally hit in the vehicle make is used to assess the (non-)safety of a make. Make sure to handle the missing values appropriately. (look up variable MAKE)

```{r}
col <- pers[, c("INJ_SEV", "MAKE", "STATE", "COUNTY")]
danger <- col %>%
  filter(INJ_SEV == 4) %>%
  group_by(STATE, MAKE) %>%
  summarize(n = n()) %>%
  group_by(STATE) %>%
  filter(n == max(n)) %>%
  filter(is.na(MAKE) == FALSE) %>%
  arrange(desc(n))
danger

```

4)Create a map, and label each state with the most dangerous vehicle. Discuss the definition of the most dangerous vehicle, and what you find from the map. (Hint: Read the description for the STATE and COUNTY columns in the FARS manual. The state & county codes are Geographic Locator Codes (GLCs) from the General Services Administration’s (GSA) publication. Use readxl::read_xlsx to read in the GLCs.)
```{r}

codes<- readxl::read_xlsx("codes.xlsx")
makeCodes<- read.csv("makeCodes.csv")
stateCodes<- read.csv('stateCodes.csv')
colnames(stateCodes)[1] <- "STATE"
colnames(stateCodes)[2] <- "State Name"


tempDF<-col %>% 
  group_by(STATE,COUNTY, MAKE) %>% 
  summarize(n = n()) %>% 
   group_by(STATE) %>%
  filter(is.na(MAKE) == FALSE) %>%
  filter(n == max(n)) %>% 
  arrange(desc(n)) %>% 
  rename("Codes" = "MAKE")

mainDF<- left_join(tempDF, makeCodes, by ="Codes")

mainDF<-left_join(mainDF, stateCodes, by ="STATE")

#Get the average Lat and Long to position the labels nicely later
spatialDF <- acc[,c("STATE","COUNTY","LATITUDE", "LONGITUD")]
spatialDF<- spatialDF %>% group_by(STATE,COUNTY) %>% summarise(aveLong = (min(LONGITUD) + max(LONGITUD))/2, aveLat = (min(LATITUDE) + max(LATITUDE))/2)
finalDF<- left_join(mainDF, spatialDF, by = c("STATE", "COUNTY"))

states <- map_data('state')


ggplot(states, aes(x=long, y=lat)) + geom_polygon(aes(group = group), fill = "white", colour = "grey50") + geom_text(data = finalDF, aes(x=aveLong, y= aveLat, label = Make)) +coord_cartesian(xlim=c(-130,-60), ylim =c(20,50))



# codes <- codes[, c( "State Code", "County Code", "State Name","County Name")] %>%
#   rename("STATE" = "State Code") %>% 
#   rename("COUNTY" = "County Code")

##codes
# codes$STATE <- as.numeric(codes$STATE)
# codes$COUNTY <- as.numeric(codes$COUNTY)
# 
# #Main DataFrames for Map
# 
# masterDF<-full_join(codes, spatialDF, by = "COUNTY","STATE")
# masterDF<-masterDF %>% rename("STATE" = "STATE.x")

#spatialDF$STATE <- as.numeric(spatialDF$STATE)
#spatialDF$COUNTY <- as.numeric(spatialDF$COUNTY)






#masterDF<-full_join(masterDF, tempDF,by = "COUNTY","STATE")
#masterDF %>% drop_na()
#Choropleth Map

```

5)
```{r}

joinedTables <- right_join(pers,acc)

```
6)
```{r}
joinedTables <- joinedTables %>%
  group_by(DAY_WEEK) %>%
  group_by(HOUR) %>%
  group_by(SEX)
head(joinedTables)

#Male Datasets & Plots
male <- subset(joinedTables, SEX == 1)

m_hour<- male %>% 
  group_by(HOUR) %>% 
  summarise(n =n()) 
ggplot(m_hour, aes(x=HOUR, y =n))+ geom_bar(stat = "identity")+coord_cartesian(xlim=c(0,25))+ labs(x="Hour of Day", y = "Number of male Accidents",
       title = "Accidents vs Hour of the Day (MALE)")

m_day<- male %>% 
  group_by(DAY_WEEK) %>% 
  summarise(n =n())
ggplot(m_day, aes(x=DAY_WEEK, y =n))+ geom_bar(stat = "identity") + labs(x="Day of the Week", y = "Number of male Accidents",
       title = "Accidents vs Day of the Week (MALE)")
  
#Female Datasets & Plots
female <- subset(joinedTables, SEX == 2)
f_hour<- female %>% 
  group_by(HOUR) %>% 
  summarise(n =n())
ggplot(f_hour, aes(x=HOUR, y =n))+ geom_bar(stat = "identity")+coord_cartesian(xlim=c(0,24)) + labs(x="Hour of Day", y = "Number of female Accidents",
       title = "Accidents vs Hour of the Day (FEMALE)")


f_day<- female %>% 
  group_by(DAY_WEEK) %>% 
  summarise(n =n())
ggplot(f_day, aes(x=DAY_WEEK, y =n))+ geom_bar(stat = "identity") + labs(x="Day of the Week", y = "Number of female Accidents",
       title = "Accidents vs Day of the Week (FEMALE)")
  



```
