---
title: "team"
author: "Rabsa Naseer & Ahmed Hassan"
date: "11/9/2020"
output: html_document
---

```{r}
library(ggplot2)
library(readxl)
library(dplyr)
library(tidyverse)
library(lubridate)
library(ggrepel)

acc <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/accident.csv", stringsAsFactors = FALSE)
pers  <- read.csv("https://raw.githubusercontent.com/xdaiISU/ds202materials/master/hwlabs/fars2017/person.csv", stringsAsFactors = FALSE)
```

```{r}
make <- readr::read_csv("makeCodes.xls")%>%
  rename("STATE" = "Codes")
state <- readr::read_csv("stateCodes.xls") %>%
  rename("STATE" = "Codes")

##adding state and make label columns
acc2 <- right_join(state, acc)
acc_final <- right_join(make, acc2)
pers2 <- right_join(make, pers)
pers_final <- right_join(state, pers2)
acc_final
day <- readr::read_csv("dayCodes.csv")

```

1.
```{r}
##Adding day names column
acc_days <-merge(day, acc_final, by="DAY_WEEK")
acc_days %>% group_by(Day) %>% 
  summarise(total = length(Day)) %>%
  arrange(desc(total))
```
According to the data, the weekend has the most accidents, Saturday with the highest number of accidents. Friday and Sunday have similar numbers of accidents.

2.
```{r}
hurt <- pers_final[pers_final$INJ_SEV == 4,]
```

3.
```{r}
col <- pers_final[, c("INJ_SEV", "MAKE", "Make", "STATE","State", "COUNTY")]
danger <- col %>%
  filter(INJ_SEV == 4) %>%
  group_by(STATE, State, MAKE, Make) %>%
  summarize(n = n()) %>%
  group_by(STATE) %>%
  filter(n == max(n)) %>%
  filter(is.na(MAKE) == FALSE) %>%
  arrange(desc(n))
head(danger)
```

4.
```{r}

codes<- readxl::read_xlsx("codes.xlsx")
makeCodes<- read.csv("makeCodes.csv")
stateCodes<- read.csv('stateCodes.csv')
colnames(stateCodes)[1] <- "STATE"
colnames(stateCodes)[2] <- "State Name"


tempDF<-col %>% 
  group_by(STATE,COUNTY, MAKE) %>% 
  summarize(n = n()) %>% 
   group_by(STATE) %>%
  filter(is.na(MAKE) == FALSE) %>%
  filter(n == max(n)) %>% 
  arrange(desc(n)) %>% 
  rename("Codes" = "MAKE")

mainDF<- left_join(tempDF, makeCodes, by ="Codes")

mainDF<-left_join(mainDF, stateCodes, by ="STATE")

#Get the average Lat and Long to position the labels nicely later
spatialDF <- acc[,c("STATE","COUNTY","LATITUDE", "LONGITUD")]
spatialDF<- spatialDF %>% group_by(STATE,COUNTY) %>% summarise(aveLong = (min(LONGITUD) + max(LONGITUD))/2, aveLat = (min(LATITUDE) + max(LATITUDE))/2)
finalDF<- left_join(mainDF, spatialDF, by = c("STATE", "COUNTY"))

states <- map_data('state')


ggplot(states, aes(x=long, y=lat)) + geom_polygon(aes(group = group), fill = "white", colour = "grey50") + geom_text(data = finalDF, aes(x=aveLong, y= aveLat, label = Make)) +coord_cartesian(xlim=c(-130,-60), ylim =c(20,50))





```


5.
```{r}
joinedTables <- right_join(pers_final,acc_final)
```


6.
```{r}
joinedTables <- joinedTables %>%
  group_by(DAY_WEEK) %>%
  group_by(HOUR) %>%
  group_by(SEX)
head(joinedTables)

#Male Datasets & Plots
male <- subset(joinedTables, SEX == 1)

m_hour<- male %>% 
  group_by(HOUR) %>% 
  summarise(n =n()) 
ggplot(m_hour, aes(x=HOUR, y =n))+ geom_bar(stat = "identity")+coord_cartesian(xlim=c(0,25))+ labs(x="Hour of Day", y = "Number of male Accidents",
       title = "Accidents vs Hour of the Day (MALE)")

m_day<- male %>% 
  group_by(DAY_WEEK) %>% 
  summarise(n =n())
ggplot(m_day, aes(x=DAY_WEEK, y =n))+ geom_bar(stat = "identity") + labs(x="Day of the Week", y = "Number of male Accidents",
       title = "Accidents vs Day of the Week (MALE)")
  
#Female Datasets & Plots
female <- subset(joinedTables, SEX == 2)
f_hour<- female %>% 
  group_by(HOUR) %>% 
  summarise(n =n())
ggplot(f_hour, aes(x=HOUR, y =n))+ geom_bar(stat = "identity")+coord_cartesian(xlim=c(0,24)) + labs(x="Hour of Day", y = "Number of female Accidents",
       title = "Accidents vs Hour of the Day (FEMALE)")


f_day<- female %>% 
  group_by(DAY_WEEK) %>% 
  summarise(n =n())
ggplot(f_day, aes(x=DAY_WEEK, y =n))+ geom_bar(stat = "identity") + labs(x="Day of the Week", y = "Number of female Accidents",
       title = "Accidents vs Day of the Week (FEMALE)")
  



```


7.
```{r}
joinedTables %>% filter(INJ_SEV ==4)
sevenDF<- joinedTables[,c("STATE","COUNTY","DOA")]
sevenDF<- sevenDF %>% group_by(STATE,COUNTY) %>%
  summarise(sumCount = sum(DOA))

county <- map_data('county')
county <-county[,c("group","order","long","lat")]
colnames(county)[1] <- "STATE"
colnames(county)[2] <- "COUNTY"

lastDF<- right_join(sevenDF,county, by= "STATE")
colnames(lastDF)[2] <- "COUNTY"
lastDF$sumCount[is.na(lastDF$sumCount)] <- 0

# 
# lastDF %>% group_by(STATE,COUNTY) %>% 
#     summarise(sumCount = sum(DOA))


#make plot
ggplot(lastDF, aes(x = long, y = lat, fill = lastDF$sumCount)) + 
  geom_polygon() + 
  geom_path(data = states, aes(x = long, y = lat, group = group), inherit.aes=FALSE, size=2, color='white')



```

8.
```{r}
summer <- joinedTables %>%
  filter((DAY >= 20 & MONTH >= 6) & (DAY <= 22 & MONTH <= 9)) %>%
  filter(INJ_SEV == 4 | 3) %>%
  group_by(STATE, State) %>%
  summarize(n = n()) %>%
  group_by(STATE) %>%
  filter(n == max(n)) %>%
  arrange(desc(n))
head(summer)
winter <- joinedTables %>%
  filter((DAY >= 21 & MONTH == 12) | (DAY >= 1 & MONTH >= 1) & (DAY <= 1 & MONTH <= 3)) %>%
  filter(INJ_SEV == 4 | 3) %>%
  group_by(STATE, State) %>%
  summarize(n = n()) %>%
  group_by(STATE) %>%
  filter(n == max(n)) %>%
  arrange(desc(n))
head(winter)
sums <- data.frame("Summer" = sum(summer$n), "Winter" = sum(winter$n))
sums
```
Winter seems to have more dangerous than summer when considering the number of severe accidents and suspected serious injury accidents in total. The number of these two types of accidents in each state differ, but it seems in both seasons that larger states or states with higher populations have higher frequencies.
